### 2022-11-10 submission/manuscript.tex
Change coverage from fraction to percent

\begin{itemize}
\tightlist
\item
  Rarefy your data
\item
  Problems with recommended methods\ldots{}

  \begin{itemize}
  \tightlist
  \item
    Many recommended methods are borrowed from gened expression analysis
  \item
    Meaning of zeroes in data - structural vs.~below limit of detection
  \end{itemize}
\item
  Factors that determine what number of sequences to rarefy to
\item
  Need better methods of pooling libraries that result in more even
  distribution of sequences across samples
\item
  Rarefy your data
\end{itemize}

@@

\textbf{Choice of datasets.} The specific studies used in this study
were selected because their data was publicly accessible through the
Sequence Read Archive, the original investigators sequenced the V4
region of the 16S rRNA gene using paired 250 nt reads, and my previous
familiarity with the data. The use of paired 250 nt reads to sequence
the V4 region resulted in a near complete two-fold overap of the V4
region resulting in high quality contigs with a low sequencing error
rate (1). These data were processed through a standardized sequence
curation pipeline to generate operational taxonomic units (OTUs) using
the mothur software package (1, 2). OTUs were identified using the
OptiClust algorithm to cluster sequences together that were not more
than 3\% different from each other (3).

\textbf{Null community model.} Null community models were generated such
that within a dataset the number of sequences per sample and the number
of sequences per OTU across all samples within the dataset were the same
as was observed in original. This model effectively generated
statistical samples of a single community so that there should not have
been a statistical difference between the samples. This model
implemented by randomly assigning each sequence in the dataset to an OTU
and sample while keeping constant the number of sequences per sample and
the total number of sequences in each OTU. This is a similar approach to
that of the IS algorithm described by Ulrich and Gotelli (4). Because
the construction of the null models was a stochastic process, 100
replicates were geneated for each dataset.

@@

\textbf{Test of statistical significance.} Statistical comparisons of
alpha diversity metrics across the simulated treatment groups were
performed using the non-parametric two-sample Wilcoxon test as
implemented in R. This test was selected because the alpha-diversity
metrics tended to not be normally distributed and each dataset required
a different transformation to normalize the data. Comparisons of beta
diversity metrics were performed using the adonis2 function from the
vegan (v.2.6.2) R package (5). The adonis2 function implements a
non-parametric multivariate analysis of variance using distances
matrices as described by McArdle and Anderson (6). Throughout this study
I used 0.05 as the threshold for asssessing statistical significance.

@@

\textbf{Alpha diversity calculations.} Various strategies for handling
uneven sampling effort were evaluated to identify the best approach for
calculating community richness and Shannon and inverse Simpson diversity
indeices. Raw OTU counts were used as input to calculate sample richness
and Shannon and inverse Simpson diversity using mothur (2, 7). Shannon
diversity was calculated as

@@

The inverse Simpson diversity was calculated as \(1/D_{simpson}\). In
both formulae, \(n_i\) is the number of sequences in OTU \(i\) and \(N\)
is the number of sequences in the sample. Rarefaction of richness,
Shannon diversity and Inverse Simpson diversity values were carried out
in mothur. Briefly, mothur calculates each value on a random draw of the
same number of sequences from each sample and obtains a mean value based
on 1,000 random draws. Scaled ranked subsampling (SRS) was used to
normalize OTU counts to the size of the smallest sample in each dataset
using the SRS R package (v.0.2.3)(8). Normalized OTU counts were used to
calculate sample richness and Shannon and inverse Simpson diversity
values using mothur. Data normalized by cumulative sum scaling (CSS)
were not reported for alpha-diversity values since the relative
abundances of the features do not change with the normalization
procedure (9). The non-parametric bias-corrected Chao1 and ACE richness
estimators (10) and a non-parametric estimator of the Shannon diversity
(11) were calculated using raw OTU counts with mothur. Parametric
estimates of sample richness were calculated using the breakaway (BA) R
package (v.4.7.9)(12). The current analysis reports both the results
from running default model selection procedure and the Poisson model.
The default model selection returned either the Kemp, Negative Binomial,
or Poisson models. Relative abundance data were not used to calculate
alpha diversity metrics since the richness and evenness does not change
from the raw data when dividing each sample by the total number of
sequences in the sample.

\textbf{Beta diversity calculations.} Similar to the alpha diversity
calculations, multiple approaches were used to control for uneven
sampling effort and calculate beta diversity. Raw and OTU counts were
used for input to calculate the Jaccard, Bray-Curtis, and Euclidean
dissimilarity indices using the vegdist function from the vegan R
package (v.2.6.2)(5). The Jaccard index was calculated as

@@

Three approaches were taken to normalize the number of sequences across
samples within a dataset. Scaled ranked subsampling (SRS) and cumulative
sum scaling (CSS) were used to normalize raw OTU counts using the SRS
(v.0.2.3) and metagenomeSeq (v.1.36.0) R packages, respectively (8, 9).
The normalized counts were then used to calculate Jaccard and
Bray-Curtis dissimilarity indices. Finally, the variance-stabilization
transformation (VST) as implemented in the DESeq2 (v.1.34.0) R package
was used to normalize the data as described by McMurdie and Holmes (13,
14). Because the VST approach generates negative values, which are
incompatible with calculating Jaccard and Bray-Curtis dissimilarity
values, Euclidean distances were calcualted instead.

@@

where \$x\_\{ij\} was the number of sequences observed for OTU \(i\) in
sample \(j\) and \(g()\) was the geometric mean \(x_j\) was the counts
of the \(S_T\) OTUs in sample \(j\). Because the geometric mean is zero
if any OTU is absent from a sample, the CLR is undefined when there are
unobserved OTUs in a sample. To overcome this problem, I attempted a
four approaches. The first, Zero CLR, imputed the value of the zeroes
based on the observed data using the zCompositions (v.1.4.0.1) R package
(15). The second, One CLR, added a pseudocount of 1 to the abundance of
all OTUs (9, 16). The third, Nudge CLR, added a pseudocount of 1 divided
by the total number of sequences in a sample to each OTU in the sample
(16, 17). The final approach, Robust CLR, removed unobserved OTUs prior
to calcualting the CLR (18).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-Kozich2013}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter
NT}, \textbf{Highlander SK}, \textbf{Schloss PD}. 2013. {Development of
a dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform}.
Applied and environmental microbiology \textbf{79}:5112--5120.}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2009}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Schloss PD}, \textbf{Westcott SL},
\textbf{Ryabin T}, \textbf{Hall JR}, \textbf{Hartmann M},
\textbf{Hollister EB}, \textbf{Lesniewski RA}, \textbf{Oakley BB},
\textbf{Parks DH}, \textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres
B}, \textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.}

\leavevmode\vadjust pre{\hypertarget{ref-Westcott2017}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{\textbf{Westcott SL}, \textbf{Schloss PD}. 2017.
{OptiClust}, an improved method for assigning amplicon-based sequence
data to operational taxonomic units. {mSphere} \textbf{2}.
doi:\href{https://doi.org/10.1128/mspheredirect.00073-17}{10.1128/mspheredirect.00073-17}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ulrich2010}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{\textbf{Ulrich W}, \textbf{Gotelli NJ}. 2010. Null model
analysis of species associations using abundance data. Ecology
\textbf{91}:3384--3397.
doi:\href{https://doi.org/10.1890/09-2157.1}{10.1890/09-2157.1}.}

\leavevmode\vadjust pre{\hypertarget{ref-Oksanen2022}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{\textbf{Oksanen J}, \textbf{Simpson GL},
\textbf{Blanchet FG}, \textbf{Kindt R}, \textbf{Legendre P},
\textbf{Minchin PR}, \textbf{O'Hara RB}, \textbf{Solymos P},
\textbf{Stevens MHH}, \textbf{Szoecs E}, \textbf{Wagner H},
\textbf{Barbour M}, \textbf{Bedward M}, \textbf{Bolker B},
\textbf{Borcard D}, \textbf{Carvalho G}, \textbf{Chirico M}, \textbf{De
Caceres M}, \textbf{Durand S}, \textbf{Evangelista HBA},
\textbf{FitzJohn R}, \textbf{Friendly M}, \textbf{Furneaux B},
\textbf{Hannigan G}, \textbf{Hill MO}, \textbf{Lahti L}, \textbf{McGlinn
D}, \textbf{Ouellette M-H}, \textbf{Ribeiro Cunha E}, \textbf{Smith T},
\textbf{Stier A}, \textbf{Ter Braak CJF}, \textbf{Weedon J}. 2022.
\href{https://CRAN.R-project.org/package=vegan}{Vegan: Community ecology
package}.}

\leavevmode\vadjust pre{\hypertarget{ref-McArdle2001}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{\textbf{McArdle BH}, \textbf{Anderson MJ}. 2001.
{FITTING} {MULTIVARIATE} {MODELS} {TO} {COMMUNITY} {DATA}: A {COMMENT}
{ON} {DISTANCE}-{BASED} {REDUNDANCY} {ANALYSIS}. Ecology
\textbf{82}:290--297.
doi:\href{https://doi.org/10.1890/0012-9658(2001)082\%5B0290:fmmtcd\%5D2.0.co;2}{10.1890/0012-9658(2001)082{[}0290:fmmtcd{]}2.0.co;2}.}

\leavevmode\vadjust pre{\hypertarget{ref-Magurran2004}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{\textbf{Magurran AE}. 2004.
\href{https://books.google.com/books?id=tUqzLSUzXxcC}{Measuring
biological diversity}. Wiley.}

\leavevmode\vadjust pre{\hypertarget{ref-Beule2020}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{\textbf{Beule L}, \textbf{Karlovsky P}. 2020.
\href{https://doi.org/10.7717/peerj.9593}{Improved normalization of
species count data in ecology by scaling with ranked subsampling (SRS):
Application to microbial communities}. PeerJ \textbf{8}:e9593.}

\leavevmode\vadjust pre{\hypertarget{ref-Paulson2013}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{\textbf{Paulson JN}, \textbf{Stine OC}, \textbf{Bravo
HC}, \textbf{Pop M}. 2013. Differential abundance analysis for microbial
marker-gene surveys. Nature Methods \textbf{10}:1200--1202.
doi:\href{https://doi.org/10.1038/nmeth.2658}{10.1038/nmeth.2658}.}

\leavevmode\vadjust pre{\hypertarget{ref-Chao2016}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{\textbf{Chao A}, \textbf{Chiu C-H}. 2016. Species
richness: Estimation and comparison. Wiley StatsRef: statistics
reference online \textbf{1}:26.}

\leavevmode\vadjust pre{\hypertarget{ref-Chao2003}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{\textbf{Chao A}, \textbf{Shen T-J}. 2003. Nonparametric
estimation of shannon's index of diversity when there are unseen species
in sample. Environmental and ecological statistics
\textbf{10}:429--443.}

\leavevmode\vadjust pre{\hypertarget{ref-Willis2015}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{\textbf{Willis A}, \textbf{Bunge J}. 2015. Estimating
diversity via frequency ratios. Biometrics \textbf{71}:1042--1049.}

\leavevmode\vadjust pre{\hypertarget{ref-Love2014}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{\textbf{Love MI}, \textbf{Huber W}, \textbf{Anders S}.
2014. Moderated estimation of fold change and dispersion for {RNA}-seq
data with {DESeq}2. Genome Biology \textbf{15}.
doi:\href{https://doi.org/10.1186/s13059-014-0550-8}{10.1186/s13059-014-0550-8}.}

\leavevmode\vadjust pre{\hypertarget{ref-McMurdie2014}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{\textbf{McMurdie PJ}, \textbf{Holmes S}. 2014. Waste
not, want not: Why rarefying microbiome data is inadmissible. {PLoS}
Computational Biology \textbf{10}:e1003531.
doi:\href{https://doi.org/10.1371/journal.pcbi.1003531}{10.1371/journal.pcbi.1003531}.}

\leavevmode\vadjust pre{\hypertarget{ref-Palarea2015}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{\textbf{Palarea-Albaladejo J}, \textbf{Martin-Fernandez
J}. 2015. zCompositions -- {R} package for multivariate imputation of
left-censored data under a compositional approach. Chemometrics and
Intelligent Laboratory Systems \textbf{143}:85--96.
doi:\href{https://doi.org/10.1016/j.chemolab.2015.02.019}{10.1016/j.chemolab.2015.02.019}.}

\leavevmode\vadjust pre{\hypertarget{ref-Lin2020}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{\textbf{Lin H}, \textbf{Peddada SD}. 2020. Analysis of
microbial compositions: A review of normalization and differential
abundance analysis. npj Biofilms and Microbiomes \textbf{6}.
doi:\href{https://doi.org/10.1038/s41522-020-00160-w}{10.1038/s41522-020-00160-w}.}

\leavevmode\vadjust pre{\hypertarget{ref-Costea2014}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{\textbf{Costea PI}, \textbf{Zeller G}, \textbf{Sunagawa
S}, \textbf{Bork P}. 2014. A fair comparison. Nature Methods
\textbf{11}:359--359.
doi:\href{https://doi.org/10.1038/nmeth.2897}{10.1038/nmeth.2897}.}

\leavevmode\vadjust pre{\hypertarget{ref-Martino2019}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{\textbf{Martino C}, \textbf{Morton JT}, \textbf{Marotz
CA}, \textbf{Thompson LR}, \textbf{Tripathi A}, \textbf{Knight R},
\textbf{Zengler K}. 2019. A novel sparse compositional technique reveals
microbial perturbations. {mSystems} \textbf{4}.
doi:\href{https://doi.org/10.1128/msystems.00016-19}{10.1128/msystems.00016-19}.}

>>>>

Over the past decade, the question of whether to rarefy microbial
community sequence data has been controversial. The analyses I presented
here strongly indicate that rarefaction is necessary to control for
uneven sampling effort when comparing communities based on alpha and
beta diversity indices. Compared to all other methods, rarefaction
removed the correlation between sequencing depth and alpha or beta
diversity metrics when the sequencing depth varied by as much as 97-fold
across samples. I showed that this correlation could lead falsely
detecting differences between treatment groups if sampling depth and
sequencing effort are confounded (e.g., 1). The correlation with
sampling effort leads to an artificial increase in the variation between
samples and a reduced power to detect true differences in alpha and beta
diversity. For these reasons, rarefaction is a valuable tool to control
for uneven sampling effort until improved statistical procedures are
developed or it becomes possible to more evenly distribute sequencing
effort across samples.

The primary alternative to rarefaction that many advise is to estimate
alpha diversity has been to use non-parametric or parametric methods
using raw data and to then compare the estimates (2). My data
demonstrates that such approaches are limited for several reasons.
First, it has been widely observed that non-parametric richness esimates
such as ACE and Chao1 are sensitive to sampling effort. Therefore, these
strategies do not, in practice, remove the effects of sampling effort.
Second, parametric approaches, such as those implemented in the
breakaway R package, generate confidence intervals that are likely to
include the true richness and theoretically shrink with increased
sampling effort. For most samples, the confidence intervals around the
estimates are too wide to be meaningful, again leading to an inability
to remvoe the effects of sampling effort. Third, it has become an
increasingly common practice for researchers to remove sequences that
are rare in a sample (e.g., those that appear once). Although that
approach was not taken in this study, removing rare sequences would skew
the distribtion of sequences and OTUs leading to a distortion of the
measurement of alpha diversity (3). The effects of removing rare
sequences would vary across samples depending on the number of sequences
in each sample. One interesting result of this analysis was the
demonstration that as metrics that depend less on rare taxa are used,
the effect of uneven sampling effort was reduced. For example, richness
counts a sequence appearing once as much as sequence appearing 1,000
times, while Shannon diversityindex places more emphasis on the more
abundant sequence, and the inverse Simpson index even more. Although
normalizing communities to a common number of sequences is also
suggested (e.g., SRS normalization) to control for uneven sampling
effort, the current analysis indicates that its performance does not
meet that of rarefaction. For analysis of alpha diversity metrics,
rarefaction is the most effective and consistent approach to control fo
uneven sampling effort.

Use of relative abundances, normalized counts, variance stabilizing
transformations, and centered log ratios have each been recommended as
more robust alternatives to rarefaction. Again, the only approach that
consistently removed the effects of uneven sampling effort was
rarefaction. Most of the recommendations borrow techniques from methods
for identifying differential gene expression. Unfortunately, there is an
important but underappreciated difference between gene expression and
community data. This is the interpretation of unobserved data. For gene
expression analysis in a single organism the lack of any sequencing data
for a gene would indicate that its expression was below the limit of
detection. Sequencing the same organism under multiple conditions would
not lead to a seemingly unbouned number of genes in the organims.
Rather, the number of genes has a definite limit that is knowable from
the genome sequence. With microbiome data, an unobserved sequence could
mean that the organism was present, but below the limit of detection or
that the organism was missing. Because we have yet to exhaustively
sample any community in the same way we have sequenced a single genome
it is unreliable to impute the presence of all organisms. Yet, this is
exactly what the variance stabilization transformation and most CLR
techniques do. This analysis has demonstrated a clear correlation
between distances calculated by these methods and sequencing effort.
This result is at odds with the claims by others that the distances are
scale invariant (4). Again, rarefaction is the most effective and
consistent approach to control for uneven sampling effort when
calculating beta diversity metrics.

Two common critique of rarefaction is that the approach ommits valid
data and that the selection of the value to rarefy the sequence data to
is arbitrary (5). I disagree with both critiques. All of the data are
used to calculate the mean value of the rarefied metrics; none of it is
excluded. When the dataset is subsampled, every sequence has a random
chance of being included in the calculated metric. When that subsampling
is repeated a large number of times (e.g., 100 or 1000) the risk of
oversampling rare taxa is mititgated. As for the second critism, I would
resist the claim that the selection of the rarefaction threshold is
arbitrary. In practice, there is a tradeoff between sampling breadth and
depth. Greater breadth will increase the statistical power to compare
treatment groups and greater depth will increase the resolution to
describe the communities. My process for picking a break involves
looking for a natural break in the distribution of the number of
sequences. For example, the Lake dataset used in this study had a clear
break around 10,000 sequences. I would also consider what samples are
below any break that I select. If there were critical samples below the
break I would either reduce the break point or I would generate more
sequences from those samples. Furthermore, regardless of the break point
one selects, they could repeat an analysis at multiple thresholds to
assess the tradeoff between statistical power and resolution. As shown
in my analysis of Good's coverage values, most studies obtain an ample
level of coverage and would need to increase their sampling depth by 10
fold to increase the coverage by several percentage points. I contend
that for most studies increased sequencing effort should be applied to
improve sampling breadth and statistica power over depth and resolution.

\begin{itemize}
\tightlist
\item
  Need better methods of pooling libraries that result in more even
  distribution of sequences across samples
\item
  We don't have true absolute numbers
\item
  Rarefy your data
\end{itemize}

@@

\textbf{Choice of datasets.} The specific studies used in this study
were selected because their data was publicly accessible through the
Sequence Read Archive, the original investigators sequenced the V4
region of the 16S rRNA gene using paired 250 nt reads, and my previous
familiarity with the data. The use of paired 250 nt reads to sequence
the V4 region resulted in a near complete two-fold overap of the V4
region resulting in high quality contigs with a low sequencing error
rate (6). These data were processed through a standardized sequence
curation pipeline to generate operational taxonomic units (OTUs) using
the mothur software package (6, 7). OTUs were identified using the
OptiClust algorithm to cluster sequences together that were not more
than 3\% different from each other (8).

\textbf{Null community model.} Null community models were generated such
that within a dataset the number of sequences per sample and the number
of sequences per OTU across all samples within the dataset were the same
as was observed in original. This model effectively generated
statistical samples of a single community so that there should not have
been a statistical difference between the samples. This model
implemented by randomly assigning each sequence in the dataset to an OTU
and sample while keeping constant the number of sequences per sample and
the total number of sequences in each OTU. This is a similar approach to
that of the IS algorithm described by Ulrich and Gotelli (9). Because
the construction of the null models was a stochastic process, 100
replicates were geneated for each dataset.

@@

\textbf{Test of statistical significance.} Statistical comparisons of
alpha diversity metrics across the simulated treatment groups were
performed using the non-parametric two-sample Wilcoxon test as
implemented in R. This test was selected because the alpha-diversity
metrics tended to not be normally distributed and each dataset required
a different transformation to normalize the data. Comparisons of beta
diversity metrics were performed using the adonis2 function from the
vegan (v.2.6.2) R package (10). The adonis2 function implements a
non-parametric multivariate analysis of variance using distances
matrices as described by McArdle and Anderson (11). Throughout this
study I used 0.05 as the threshold for asssessing statistical
significance.

@@

\textbf{Alpha diversity calculations.} Various strategies for handling
uneven sampling effort were evaluated to identify the best approach for
calculating community richness and Shannon and inverse Simpson diversity
indeices. Raw OTU counts were used as input to calculate sample richness
and Shannon and inverse Simpson diversity using mothur (7, 12). Shannon
diversity was calculated as

@@

The inverse Simpson diversity was calculated as \(1/D_{simpson}\). In
both formulae, \(n_i\) is the number of sequences in OTU \(i\) and \(N\)
is the number of sequences in the sample. Rarefaction of richness,
Shannon diversity and Inverse Simpson diversity values were carried out
in mothur. Briefly, mothur calculates each value on a random draw of the
same number of sequences from each sample and obtains a mean value based
on 1,000 random draws. Scaled ranked subsampling (SRS) was used to
normalize OTU counts to the size of the smallest sample in each dataset
using the SRS R package (v.0.2.3)(13). Normalized OTU counts were used
to calculate sample richness and Shannon and inverse Simpson diversity
values using mothur. Data normalized by cumulative sum scaling (CSS)
were not reported for alpha-diversity values since the relative
abundances of the features do not change with the normalization
procedure (14). The non-parametric bias-corrected Chao1 and ACE richness
estimators (15) and a non-parametric estimator of the Shannon diversity
(16) were calculated using raw OTU counts with mothur. Parametric
estimates of sample richness were calculated using the breakaway (BA) R
package (v.4.7.9)(17). The current analysis reports both the results
from running default model selection procedure and the Poisson model.
The default model selection returned either the Kemp, Negative Binomial,
or Poisson models. Relative abundance data were not used to calculate
alpha diversity metrics since the richness and evenness does not change
from the raw data when dividing each sample by the total number of
sequences in the sample.

\textbf{Beta diversity calculations.} Similar to the alpha diversity
calculations, multiple approaches were used to control for uneven
sampling effort and calculate beta diversity. Raw and OTU counts were
used for input to calculate the Jaccard, Bray-Curtis, and Euclidean
dissimilarity indices using the vegdist function from the vegan R
package (v.2.6.2)(10). The Jaccard index was calculated as

@@

Three approaches were taken to normalize the number of sequences across
samples within a dataset. Scaled ranked subsampling (SRS) and cumulative
sum scaling (CSS) were used to normalize raw OTU counts using the SRS
(v.0.2.3) and metagenomeSeq (v.1.36.0) R packages, respectively (13,
14). The normalized counts were then used to calculate Jaccard and
Bray-Curtis dissimilarity indices. Finally, the variance-stabilization
transformation (VST) as implemented in the DESeq2 (v.1.34.0) R package
was used to normalize the data as described by McMurdie and Holmes (5,
18). Because the VST approach generates negative values, which are
incompatible with calculating Jaccard and Bray-Curtis dissimilarity
values, Euclidean distances were calcualted instead.

@@

where \$x\_\{ij\} was the number of sequences observed for OTU \(i\) in
sample \(j\) and \(g()\) was the geometric mean \(x_j\) was the counts
of the \(S_T\) OTUs in sample \(j\). Because the geometric mean is zero
if any OTU is absent from a sample, the CLR is undefined when there are
unobserved OTUs in a sample. To overcome this problem, I attempted a
four approaches. The first, Zero CLR, imputed the value of the zeroes
based on the observed data using the zCompositions (v.1.4.0.1) R package
(19). The second, One CLR, added a pseudocount of 1 to the abundance of
all OTUs (14, 20). The third, Nudge CLR, added a pseudocount of 1
divided by the total number of sequences in a sample to each OTU in the
sample (20, 21). The final approach, Robust CLR, removed unobserved OTUs
prior to calcualting the CLR (4).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-Morris2013}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Morris A}, \textbf{Beck JM}, \textbf{Schloss
PD}, \textbf{Campbell TB}, \textbf{Crothers K}, \textbf{Curtis JL},
\textbf{Flores SC}, \textbf{Fontenot AP}, \textbf{Ghedin E},
\textbf{Huang L}, \textbf{Jablonski K}, \textbf{Kleerup E},
\textbf{Lynch SV}, \textbf{Sodergren E}, \textbf{Twigg H}, \textbf{Young
VB}, \textbf{Bassis CM}, \textbf{Venkataraman A}, \textbf{Schmidt TM},
\textbf{Weinstock GM}. 2013. Comparison of the respiratory microbiome in
healthy nonsmokers and smokers. American Journal of Respiratory and
Critical Care Medicine \textbf{187}:1067--1075.
doi:\href{https://doi.org/10.1164/rccm.201210-1913oc}{10.1164/rccm.201210-1913oc}.}

\leavevmode\vadjust pre{\hypertarget{ref-Willis2019}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Willis AD}. 2019. Rarefaction, alpha diversity,
and statistics. Frontiers in Microbiology \textbf{10}.
doi:\href{https://doi.org/10.3389/fmicb.2019.02407}{10.3389/fmicb.2019.02407}.}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2020}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{\textbf{Schloss PD}. 2020. Removal of rare amplicon
sequence variants from 16S {rRNA} gene sequence surveys biases the
interpretation of community structure data.
doi:\href{https://doi.org/10.1101/2020.12.11.422279}{10.1101/2020.12.11.422279}.}

\leavevmode\vadjust pre{\hypertarget{ref-Martino2019}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{\textbf{Martino C}, \textbf{Morton JT}, \textbf{Marotz
CA}, \textbf{Thompson LR}, \textbf{Tripathi A}, \textbf{Knight R},
\textbf{Zengler K}. 2019. A novel sparse compositional technique reveals
microbial perturbations. {mSystems} \textbf{4}.
doi:\href{https://doi.org/10.1128/msystems.00016-19}{10.1128/msystems.00016-19}.}

\leavevmode\vadjust pre{\hypertarget{ref-McMurdie2014}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{\textbf{McMurdie PJ}, \textbf{Holmes S}. 2014. Waste
not, want not: Why rarefying microbiome data is inadmissible. {PLoS}
Computational Biology \textbf{10}:e1003531.
doi:\href{https://doi.org/10.1371/journal.pcbi.1003531}{10.1371/journal.pcbi.1003531}.}

\leavevmode\vadjust pre{\hypertarget{ref-Kozich2013}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{\textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter
NT}, \textbf{Highlander SK}, \textbf{Schloss PD}. 2013. {Development of
a dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform}.
Applied and environmental microbiology \textbf{79}:5112--5120.}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2009}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{\textbf{Schloss PD}, \textbf{Westcott SL},
\textbf{Ryabin T}, \textbf{Hall JR}, \textbf{Hartmann M},
\textbf{Hollister EB}, \textbf{Lesniewski RA}, \textbf{Oakley BB},
\textbf{Parks DH}, \textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres
B}, \textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.}

\leavevmode\vadjust pre{\hypertarget{ref-Westcott2017}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{\textbf{Westcott SL}, \textbf{Schloss PD}. 2017.
{OptiClust}, an improved method for assigning amplicon-based sequence
data to operational taxonomic units. {mSphere} \textbf{2}.
doi:\href{https://doi.org/10.1128/mspheredirect.00073-17}{10.1128/mspheredirect.00073-17}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ulrich2010}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{\textbf{Ulrich W}, \textbf{Gotelli NJ}. 2010. Null model
analysis of species associations using abundance data. Ecology
\textbf{91}:3384--3397.
doi:\href{https://doi.org/10.1890/09-2157.1}{10.1890/09-2157.1}.}

\leavevmode\vadjust pre{\hypertarget{ref-Oksanen2022}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{\textbf{Oksanen J}, \textbf{Simpson GL},
\textbf{Blanchet FG}, \textbf{Kindt R}, \textbf{Legendre P},
\textbf{Minchin PR}, \textbf{O'Hara RB}, \textbf{Solymos P},
\textbf{Stevens MHH}, \textbf{Szoecs E}, \textbf{Wagner H},
\textbf{Barbour M}, \textbf{Bedward M}, \textbf{Bolker B},
\textbf{Borcard D}, \textbf{Carvalho G}, \textbf{Chirico M}, \textbf{De
Caceres M}, \textbf{Durand S}, \textbf{Evangelista HBA},
\textbf{FitzJohn R}, \textbf{Friendly M}, \textbf{Furneaux B},
\textbf{Hannigan G}, \textbf{Hill MO}, \textbf{Lahti L}, \textbf{McGlinn
D}, \textbf{Ouellette M-H}, \textbf{Ribeiro Cunha E}, \textbf{Smith T},
\textbf{Stier A}, \textbf{Ter Braak CJF}, \textbf{Weedon J}. 2022.
\href{https://CRAN.R-project.org/package=vegan}{Vegan: Community ecology
package}.}

\leavevmode\vadjust pre{\hypertarget{ref-McArdle2001}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{\textbf{McArdle BH}, \textbf{Anderson MJ}. 2001.
{FITTING} {MULTIVARIATE} {MODELS} {TO} {COMMUNITY} {DATA}: A {COMMENT}
{ON} {DISTANCE}-{BASED} {REDUNDANCY} {ANALYSIS}. Ecology
\textbf{82}:290--297.
doi:\href{https://doi.org/10.1890/0012-9658(2001)082\%5B0290:fmmtcd\%5D2.0.co;2}{10.1890/0012-9658(2001)082{[}0290:fmmtcd{]}2.0.co;2}.}

\leavevmode\vadjust pre{\hypertarget{ref-Magurran2004}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{\textbf{Magurran AE}. 2004.
\href{https://books.google.com/books?id=tUqzLSUzXxcC}{Measuring
biological diversity}. Wiley.}

\leavevmode\vadjust pre{\hypertarget{ref-Beule2020}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{\textbf{Beule L}, \textbf{Karlovsky P}. 2020.
\href{https://doi.org/10.7717/peerj.9593}{Improved normalization of
species count data in ecology by scaling with ranked subsampling (SRS):
Application to microbial communities}. PeerJ \textbf{8}:e9593.}

\leavevmode\vadjust pre{\hypertarget{ref-Paulson2013}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{\textbf{Paulson JN}, \textbf{Stine OC}, \textbf{Bravo
HC}, \textbf{Pop M}. 2013. Differential abundance analysis for microbial
marker-gene surveys. Nature Methods \textbf{10}:1200--1202.
doi:\href{https://doi.org/10.1038/nmeth.2658}{10.1038/nmeth.2658}.}

\leavevmode\vadjust pre{\hypertarget{ref-Chao2016}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{\textbf{Chao A}, \textbf{Chiu C-H}. 2016. Species
richness: Estimation and comparison. Wiley StatsRef: statistics
reference online \textbf{1}:26.}

\leavevmode\vadjust pre{\hypertarget{ref-Chao2003}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{\textbf{Chao A}, \textbf{Shen T-J}. 2003. Nonparametric
estimation of shannon's index of diversity when there are unseen species
in sample. Environmental and ecological statistics
\textbf{10}:429--443.}

\leavevmode\vadjust pre{\hypertarget{ref-Willis2015}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{\textbf{Willis A}, \textbf{Bunge J}. 2015. Estimating
diversity via frequency ratios. Biometrics \textbf{71}:1042--1049.}

\leavevmode\vadjust pre{\hypertarget{ref-Love2014}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{\textbf{Love MI}, \textbf{Huber W}, \textbf{Anders S}.
2014. Moderated estimation of fold change and dispersion for {RNA}-seq
data with {DESeq}2. Genome Biology \textbf{15}.
doi:\href{https://doi.org/10.1186/s13059-014-0550-8}{10.1186/s13059-014-0550-8}.}
